---
title: "SocialMedia vs Productivity"
output: html_notebook
---
```{r}
library(tidyverse)
library(ggplot2)
df <- read_csv("social_media_vs_productivity.csv")
head(df)
```
```{r}
summary(df)
```
```{r}
str(df)
```
```{r}
colSums(is.na(df))
```
```{r}
df <- df %>% rename(
  "job" = "job_type",
  "online_hours" = "daily_social_media_time",
  "app" = "social_platform_preference" ,
  "notifications" = "number_of_notifications",
  "work_hours" = "work_hours_per_day",
  "self_prod_score" = "perceived_productivity_score",
  "actual_prod_score" = "actual_productivity_score",
  "stress" = "stress_level",
  "focus_app" = "uses_focus_apps",
  "digital_wellbeing" = "has_digital_wellbeing_enabled",
  "coffee" = "coffee_consumption_per_day",
  "burnout_days" = "days_feeling_burnout_per_month",
  "job_score" = "job_satisfaction_score"
)
```
```{r}
cols_to_fill <- c("online_hours", "self_prod_score", "actual_prod_score", "stress", "sleep_hours", "screen_time_before_sleep", "job_score")  # Replace with your column names
for (col in cols_to_fill) {
  df[[col]][is.na(df[[col]])] <- mean(df[[col]], na.rm = TRUE)
}
colSums(is.na(df))
```
```{r}
cols_to_round <- c("self_prod_score", "actual_prod_score", "job_score")
for (col in cols_to_round) {
  df[[col]] <- round(df[[col]])
}
head(df)
```


```{r}
df <- df[df$gender != "Other", ]
ggplot(df, aes(x = `gender`)) +
  geom_bar(fill = "cyan", color = "black", alpha = 0.5 , stat = "count") +
  labs(title = "Gender Distribution", x = "Gender", y = "Frequency")
```
```{r}
ggplot(df, aes(x = `job`)) +
  geom_bar(fill = "green", color = "black", alpha = 0.5) +
  labs(title = "Job Distribution", x = "Job", y = "Frequency")
```
```{r}
ggplot(df, aes(x = `focus_app`)) +
  geom_bar(fill = "brown", color = "black", alpha = 0.5) +
  labs(title = "Focus App Users", x = "Yes or No", y = "Frequency")
```
```{r}
ggplot(df, aes(x = `digital_wellbeing`)) +
  geom_bar(fill = "red", color = "black", alpha = 0.5) +
  labs(title = "Digital Wellbeing Enabled", x = "Yes or No", y = "Frequency")
```
```{r}
ggplot(df, aes(x = `app`)) +
  geom_bar(fill = "purple", color = "black", alpha = 0.5 , stat = "count") +
  labs(title = "Social Media Platform Distribution", x = "App", y = "Frequency")
```
```{r}
ggplot(df, aes(x = `age`)) +
  geom_density(fill = "yellow", color = "black", alpha = 0.5) +
  labs(title = "Age", x = "Age", y = "Frequency")
```
```{r}
df <- df %>%
  mutate(age_group = cut(
    age,
    breaks = c(17, 30, 47, 65),
    labels = c("Young Adult", "Middle-Aged Adults", "Old Adults"),
    right = TRUE,
    include.lowest = TRUE
  ))
age_counts <- table(df$age_group)
pct <- round(100 * age_counts / sum(age_counts), 1)
labels <- paste(names(age_counts), "\n", pct, "%")
pie(age_counts, labels = labels, main = "Age Group Distribution")
```
```{r}
ggplot(df, aes(x = `self_prod_score`)) +
  geom_histogram(binwidth = 1, fill = "cyan", color = "black", alpha = 0.5) +
  scale_x_continuous(breaks = 1:10) +
  labs(title = "Self Productivity Score", x = "Self Productivity Score", y = "Frequency")
```
```{r}
ggplot(df, aes(x = `actual_prod_score`)) +
  geom_histogram(binwidth = 1, fill = "green", color = "black", alpha = 0.5) +
  scale_x_continuous(breaks = 1:10) +
  labs(title = "Actual Prodcutivity Score", x = "Actual Productivity Score", y = "Frequency")
```
```{r}
ggplot(df, aes(x = `sleep_hours`)) +
  geom_density(fill = "red", color = "black", alpha = 0.5) +
  labs(title = "Sleep Hours Distirbution", x = "Sleep Hours", y = "Frequency")
```
```{r}
ggplot(df, aes(y = `work_hours`)) +
  geom_boxplot(outlier.colour = "red", fill = "red", outlier.shape = 16) +
  labs(title = "Boxplot of Sleep Hours", y = "Hours")
```
```{r}
ggplot(df, aes(x = `work_hours`)) +
  geom_density(fill = "cyan", color = "black", alpha = 0.5) +
  labs(title = "Work Hours Distirbution", x = "Work Hours", y = "Frequency")
```
```{r}
ggplot(df, aes(y = `work_hours`)) +
  geom_boxplot(outlier.colour = "red", fill = "cyan", outlier.shape = 16) +
  labs(title = "Boxplot of Work Hours", y = "Hours")
```
```{r}
ggplot(df, aes(x = `online_hours`)) +
  geom_density(fill = "pink", color = "black", alpha = 0.5) +
  labs(title = "Online Hours Distirbution", x = "Online Hours", y = "Frequency")
```
```{r}
ggplot(df, aes(y = `online_hours`)) +
  geom_boxplot(outlier.colour = "red", fill = "pink", outlier.shape = 16) +
  labs(title = "Boxplot of Online Hours", y = "Hours")
```
```{r}
replace_outliers_with_mean <- function(data, columns) {
  for (col in columns) {
      Q1 <- quantile(data[[col]], 0.25, na.rm = TRUE)
      Q3 <- quantile(data[[col]], 0.75, na.rm = TRUE)
      IQR_val <- Q3 - Q1

      lower_bound <- Q1 - 1.5 * IQR_val
      upper_bound <- Q3 + 1.5 * IQR_val

      mean_val <- mean(data[[col]], na.rm = TRUE)

      data[[col]] <- ifelse(data[[col]] < lower_bound | data[[col]] > upper_bound,
                            mean_val,
                            data[[col]])
  }
  return(data)
}
df <- replace_outliers_with_mean(df, c("online_hours", "work_hours", "sleep_hours"))
```
```{r}
ggplot(df, aes(y = `online_hours`)) +
  geom_boxplot(outlier.colour = "red", fill = "pink", outlier.shape = 16) +
  labs(title = "Boxplot of Online Hours", y = "Hours")
```
```{r}
ggplot(df, aes(y = `work_hours`)) +
  geom_boxplot(outlier.colour = "red", fill = "cyan", outlier.shape = 16) +
  labs(title = "Boxplot of Work Hours", y = "Hours")
```
```{r}
ggplot(df, aes(y = `work_hours`)) +
  geom_boxplot(outlier.colour = "red", fill = "red", outlier.shape = 16) +
  labs(title = "Boxplot of Sleep Hours", y = "Hours")
```
```{r}
ggplot(df, aes(x = `stress`)) +
  geom_histogram(binwidth = 1, fill = "green", color = "black", alpha = 0.5) +
  scale_x_continuous(breaks = 1:10) +
  labs(title = "Stress Level", x = "Stress Level", y = "Frequency")
```
```{r}
ggplot(df, aes(x = `job_score`)) +
  geom_histogram(binwidth = 1,fill = "purple", color = "black", alpha = 0.5) +
  labs(title = "Job Satisfaction Score", x = "Job Satisfaction Score", y = "Frequency")
```
```{r}
library(ggcorrplot)
corr_mat = round(cor(df[sapply(df, is.numeric)]),1)
ggcorrplot(corr_mat, lab = TRUE)
```
No strong correlation between features in dataset so predictive analysis won't be feasible. 
so leaning towards hypothesis testing.
We will be using t-test, anova to test some basic hypothesis.

H0: Average productivity score is equal to 6
Ha: Average productivity score is not equal to 6
```{r}
t_test_productivity <- t.test(df$actual_prod_score, mu = 6)
print("T-Test Result of Actual Productivity Score")
print(t_test_productivity)
```
H0: Use of focus apps have no effect on productivity
Ha: Use of focus apps have effect on productivity
```{r}
t_test_focus_app <- t.test(actual_prod_score ~ focus_app, data = df)
print("T-Test Result of Effect of Focus Apps")
print(t_test_focus_app)
```
H0:Screen Time is equal across various Social Media platforms
Ha:Screen Time is not equal across various Social Media platforms
```{r}
anova_model <- aov(online_hours ~ app, data = df)
summary(anova_model)
```
```{r}
tukey <- TukeyHSD(anova_model)
par(mar = c(5, 9, 4, 2))
plot(tukey,las = 1)
```
H0 : Screen Time is equal across various Age group
Ha : Screen Time is not equal across various Age group
```{r}
anova_model <- aov(online_hours ~ age_group, data = df)
summary(anova_model)
```
There is no significant affect of age group in screen time.

H0: There is no difference in mean Screen Time between groups that have digital well being enabled
Ha: There is difference in mean Screen Time between groups that have digital well being enabled
```{r}
t_test_digital_wellbeing <- t.test(online_hours ~ digital_wellbeing, data = df)
print("T Test Result of Digital Wellbeing")
print(t_test_digital_wellbeing)
```